Technology for application family creation based on domain analysis

### 2016-05-04 2016-domainAnalysis/domainAnalysis.tex
added description to feature model

Then the question arise, how we need to analyze the domain for acquiring the necessary information. At present, the information gathering into knowledge bases is understood under the term ``domain analysis''. Although, Prieto-Diaz~\cite{prieto1988domain} confirms that domain analysis is an activity, which take place before system analysis and its output is used for system analysis to the same degree as system analysis’s output is used for system design.  There are other definitions of the term ``domain analysis''. Ferre~\cite{ferre1999evaluation} has presented definitions, such as: (1) the process of identification, organization and presenting the relevant information of a given domain, (2) the process, in which the customer’s knowledge are identified, concretized and systemized. The relevant information of the domain should be presented in objective, readily available way, such way is called ``domain model''. Mernik~\cite{mernik2005} specifies that the domain model includes not only glossary, but also must describe commonalities and variabilities of terms. Such model should precisely set bounds of the domain, i.e. clear and exact characterize a range of questions, which are considered in the domain. Term variabilities allow to define exactly, what information must be specified in concrete system implementation. Term commonalities are used for defining a set of shared operations between different applications. Implementing commonalities and adding the gained model  with information, which can be specified in instance of the concrete system, a set of different systems can be obtained based on one common model. In such manner, based on one domain model, the set of different systems in given domain can be implemented. Taking into account definitions above mentioned, we can conclude that domain analysis is the activity of forward system analysis, which goal is to provide the domain model.

As stated above, at present in many software companies the term ``domain analysis'' is understood as information gathering into some knowledge bases, but it is obvious that there are disadvantages of this approach. It may lead to incomplete glossary, absence of agreements about understanding some terms in the domain, so any misunderstanding of domain can result in an improper product. Therefore, several dozens of years ago were introduced some formal approaches for domain analysis. Here will be mentioned some of them. It is worth noting that main objective any domain analysis approach is to gain the domain model.

@@

As mentioned above, it was used QReal as DSM tool. A visual language was implemented there for describing feature models. Appropriate feature model for ``Joystick'' application family is proposed on Fig.~\ref{image:joystickFeatureModel}.

@@

Then with help of QReal tool a visual language was generated. Example of generated visual language is demonstrated on Fig.~\ref{image:joystickDsl}. It can be seen that in visual language editor we can specify property ``Quantity'', explicitly pointing out concrete number of pads.

@@

To tell the truth, example is quite simple for demonstrating extensive possibilities of the approach proposed above. At present there is no rigorous evaluation of the proposed process. Also, cohesive and consistent technology for creating application family based on domain analysis is not implemented yet, here we have described a concept-proof prototype. Therefore, this work requires more detailed explorations.

\section{Conclusion}
\label{chapter:conclusion}
The problem of not using domain analysis result for further generation of some entities for software development process was stated. There were considered some formal domain analysis approaches and we concluded that creation of feature diagrams is the most elegant decision for domain analysis that can be conducted by domain expert, i.e. non-programmer, maybe in collaboration with system analysts. Moreover, there was discussed one of the possible solutions, which is presented by Estublier, we specify some problems of such method. We suggested our own approach for creation application family in one domain based on domain analysis. Also, there was some evaluation of this approach, where we pointed out that this example remains many questions because of its simplicity.

>>>>

Then the question arise, how we need to analyze the domain for acquiring the necessary information. At present, the information gathering into knowledge bases is understood under the term ``domain analysis''. Although, Prieto-Diaz~\cite{prieto1988domain} confirms that domain analysis is an activity, which is held before system analysis and its output is used for system analysis to the same degree as system analysis’s output is used for system design.  There are other definitions of the term ``domain analysis''. Ferre~\cite{ferre1999evaluation} has presented definitions, such as: (1) the process of identification, organization and presenting the relevant information of a given domain, (2) the process, in which the customer’s knowledge are identified, concretized and systemized. The relevant information of the domain should be presented in objective, readily available way, such way is called ``domain model''. Mernik~\cite{mernik2005} specifies that the domain model includes not only glossary, but also must describe commonalities and variabilities of terms. Such model should precisely set bounds of the domain, i.e. clear and exact characterize a range of questions, which are considered in the domain. Term variabilities allow to define exactly, what information must be specified in concrete system implementation. Term commonalities are used for defining a set of shared operations between different applications. Implementing commonalities and adding the gained model  with information, which can be specified in instance of the concrete system, a set of different systems can be obtained based on one common model. In such manner, based on one domain model, the set of different systems in given domain can be implemented. Taking into account definitions above mentioned, we can conclude that domain analysis is the activity of forward system analysis, which goal is to provide the domain model.

As stated above, at present in many software companies the term ``domain analysis'' is understood as information gathering into some knowledge bases, but it is obvious that there are disadvantages of this approach. It may lead to incomplete glossary, absence of agreements about understanding some terms in the domain, so any misunderstanding of domain can result in an improper product. Therefore, several dozens of years ago were introduced some formal approaches for domain analysis. Here will be mentioned some of them. Main objective any domain analysis approach is to gain the domain model.

@@

As mentioned above, it was used QReal as DSM tool. A visual language was implemented there for describing feature models. Appropriate feature model for ``Joystick'' application family is proposed on Fig.~\ref{image:joystickFeatureModel}. This feature model presents explicit features, which are labeled as green, and some unite feature groups, which are labeled as blue. Type of arrow shows which feature is compulsory (shown as solid line with arrow on the end), which compulsory but there is some alternative between them (shown as dash line with an arrow on the end), and optional features, which may be implemented but not compulsory (shown as dash line with a circle on the end).

@@

Then with the aid of QReal tool a visual language was generated. Example of generated visual language is demonstrated on Fig.~\ref{image:joystickDsl}. It can be seen that in visual language editor can be specified property ``Quantity'', explicitly noting the concrete number of pads.

@@

As can be seen, example is quite simple for demonstrating extensive possibilities of the approach proposed above. At present there is no rigorous evaluation of the proposed process. Also, cohesive and consistent technology for creating application family based on domain analysis is not implemented yet, here we have described a concept-proof prototype. Therefore, this work requires more detailed explorations.

\section{Conclusion}
\label{chapter:conclusion}
The problem of not using domain analysis result for further generation of some entities for software development process was stated. There were considered some formal domain analysis approaches and we concluded that creation of feature diagrams is the most elegant decision for domain analysis that can be conducted by domain expert, i.e. non-programmer, maybe in collaboration with system analysts. Moreover, there was discussed one of the possible solutions, which is presented by Estublier, we specify some problems of such method. We suggested our own approach for creation of application family in one domain based on domain analysis. Thus, some target applications can be implemented even by non-programmers using domain-specific language with configurating features from library. Also, there was some evaluation of this approach, where we pointed out that this example remains many questions because of its simplicity.
